Precise predictions for same-sign W-bosons scattering at the LHC

### 2018-02-12 Draft/shower.tex
some more edits in shower.tex

Finally, focusing on the third jet, we conclude the list of differential observables with the rapidity and the Zeppenfeld variable defined in Eq.~\eqref{eq:Zeppenfeld}, Figures~\ref{fig:PSy3} and~\ref{fig:PSz3}.
{\bf MZ the rapidity can be dropped (check also the following discussion in case)}
In general, for observables which involve the third jet, one
can clearly see a degradation of the agreement among the various tools, because of the poorer perturbative description of these observables. The rapidity and Zeppenfeld variable $z$
are a striking example: both at LO and NLO, the tendency of {\sc Pythia8} to generate more hard and central radiation (corresponding to low values of $z$)
is clearly visible; it is interesting to notice that
this effect survives beyond the first emission, as it can be observed by comparing {\sc Powheg-no shower} with {\sc Powheg+Pythia8}. If it is true that the central enhancement
is a bit mitigated if NLO+PS tools are used (compare LO+PS and NLO+PS from {\sc MG5\_aMC+Pythia8} with the fixed-NLO prediction), however even at NLO+PS the central region
($z_{j_3}<0.5$) is cursed by huge differences between tools. Large differences, reaching a factor 2, persist also away from the central region. \\

In conclusion, the comparison of tools including matching with parton-shower clearly shows the benefits of the inclusion of NLO corrections: for most observables described
effectively at NLO accuracy differences between tools are at (or below) the $10\%$ level. {\bf MZ compare this number with scale uncertainties}
Some exceptions exist, \emph{e.g.}\ the rapidity separation of the two tagging jets, which
on the one hand clearly suggest not to rely on a single tool/parton shower, and on the other make it worth to investigate more in details the way QCD radiation is
generated {\bf MZ cite here H VBF at NNLO?}. Finally, the size of discrepancies for observables that are described at a lower perturbative accuracy, notably those related to the third jet, suggest that
experimental analyses should rely as least as possible on those observables and, in any case, use conservative estimate of the theory uncertainties. In order to improve the
description of these observables, a simulation of VBS+j at NLO accuracy, currently unavailable but within the reach of modern automated tools, is certainly desirable.

>>>>

Finally, focusing on the third jet, we conclude the list of differential observables with the rapidity and the Zeppenfeld variable defined in Eq.~\eqref{eq:Zeppenfeld}, Figures~\ref{fig:PSy3} and~\ref{fig:PSz3}.
{\bf MZ the rapidity can be dropped (check also the following discussion in case)}
In general, for observables which involve the third jet, one
can clearly see a degradation of the agreement among the various tools, because of the poorer perturbative description of these observables. The rapidity and Zeppenfeld variable $z$
are a striking example: both at LO and NLO, the tendency of {\sc Pythia8} to generate more hard and central radiation (corresponding to low values of $z$)
is clearly visible\footnote{The effect, which is related to the way {\sc Pythia8} deals with the recoil of the radiation in VBF/S-type processes,
can be mitigated by setting {\tt SpaceShower:dipoleRecoil = on} in the {\sc Pythia8} input file (this requires version $\ge8.230$). 
However, such a setting is not compatible with the matching
in {\sc MG5\_aMC}, and has other effects, though smaller, on the rapidity spectrum of the two hardest jets.}; it is interesting to notice that
this effect survives beyond the first emission, as it can be observed by comparing {\sc Powheg-no shower} with {\sc Powheg+Pythia8}. A similar
behaviour of {\sc Pythia8}
has been also observed in the study of EW production of a Z boson in association with two jets (see the recent CMS measurement, 
Ref.~\cite{Sirunyan:2017jej} Figure 12), where the experimental data seem to prefer the description by {\sc Herwig++}. 
If it is true that the central enhancement
is a bit mitigated if NLO+PS tools are used (compare LO+PS and NLO+PS from {\sc MG5\_aMC+Pythia8} with the fixed-NLO prediction), however even at NLO+PS the central region
($z_{j_3}<0.5$) is cursed by huge differences between tools. Large differences, reaching a factor 2, persist also away from the central region. \\

In conclusion, the comparison of tools including matching with parton-shower clearly shows the benefits of the inclusion of NLO corrections: for most observables described
effectively at NLO accuracy differences between tools are at (or below) the $10\%$ level. {\bf MZ compare this number with scale uncertainties}
Some exceptions exist, \emph{e.g.}\ the rapidity separation of the two tagging jets, which
on the one hand clearly suggest not to rely on a single tool/parton shower, and on the other make it worth to investigate more in details the way QCD radiation is
generated {\bf MZ cite here H VBF at NNLO?}. Finally, the size of discrepancies for observables that are described at a lower perturbative accuracy, notably those related to the third jet, suggest that
experimental analyses should rely as least as possible on those observables and, in any case, use conservative estimate of the theory 
uncertainties. On the one hand, un order to improve the
description of these observables, a simulation of VBS+j at NLO accuracy, currently unavailable but within the reach of modern 
automated tools, is certainly desirable; on the other hand, measurements of processes with similar 
color flow (EW production of a single vector boson plus jets,
 VBF, \ldots) can certainly help in order to discriminate which tools perform better in the comparison with data.
